The CMS Data Aggregation System

### 2010-05-04 doc/paper.tex
New addition
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@1505 4525493e-7705-40b1-a816-d608a930855b

% Template for Elsevier CRC journal article
% version 1.0 dated 13 October 2009

% This file (c) 2009 Elsevier Ltd.  Modifications may be freely made,
% provided the edited file is saved under a different name

% This file contains modifications for Procedia Computer Science

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% This template uses the elsarticle.cls document class and the extension package ecrc.sty
%% For full documentation on usage of elsarticle.cls, consult the documentation "elsdoc.pdf"
%% Further resources available at http://www.elsevier.com/latex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% The '1p' and 'times' class options of elsarticle are used for Elsevier CRC
\documentclass[1p,times]{elsarticle}

%% The `ecrc' package must be called to make the CRC functionality available
\usepackage{ecrc}

%% The ecrc package defines commands needed for running heads and logos.
%% For running heads, you can set the journal name, the volume, the starting page and the authors

%% set the volume if you know. Otherwise `00'
\volume{00}

%% set the starting page if not 1
\firstpage{1}

%% Give the name of the journal
\journalname{Procedia Computer Science}

%% Give the author list to appear in the running head
%% Example \runauth{C.V. Radhakrishnan et al.}
\runauth{}

%% The choice of journal logo is determined by the \jid and \jnltitlelogo commands.
%% A user-supplied logo with the name <\jid>logo.pdf will be inserted if present.
%% e.g. if \jid{yspmi} the system will look for a file yspmilogo.pdf
%% Otherwise the content of \jnltitlelogo will be set between horizontal lines as a default logo

%% Give the abbreviation of the Journal.
\jid{procs}

%% Give a short journal name for the dummy logo (if needed)
\jnltitlelogo{Procedia Computer Science}

%% Hereafter the template follows `elsarticle'.
%% For more details see the existing template files elsarticle-template-harv.tex and elsarticle-template-num.tex.

%% Elsevier CRC generally uses a numbered reference style
%% For this, the conventions of elsarticle-template-num.tex should be followed (included below)
%% If using BibTeX, use the style file elsarticle-num.bst

%% End of ecrc-specific commands
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% The amssymb package provides various useful mathematical symbols
\usepackage{amssymb}
%% The amsthm package provides extended theorem environments
%% \usepackage{amsthm}

%% The lineno packages adds line numbers. Start line numbering with
%% \begin{linenumbers}, end it with \end{linenumbers}. Or switch it on
%% for the whole article with \linenumbers after \end{frontmatter}.
%% \usepackage{lineno}

%% natbib.sty is loaded by default. However, natbib options can be
%% provided with \biboptions{...} command. Following options are
%% valid:

%%   round  -  round parentheses are used (default)
%%   square -  square brackets are used   [option]
%%   curly  -  curly braces are used      {option}
%%   angle  -  angle brackets are used    <option>
%%   semicolon  -  multiple citations separated by semi-colon
%%   colon  - same as semicolon, an earlier confusion
%%   comma  -  separated by comma
%%   numbers-  selects numerical citations
%%   super  -  numerical citations as superscripts
%%   sort   -  sorts multiple citations according to order in ref. list
%%   sort&compress   -  like sort, but also compresses numerical citations
%%   compress - compresses without sorting
%%
%% \biboptions{comma,round}

% \biboptions{}

% if you have landscape tables
\usepackage[figuresright]{rotating}

% put your own definitions here:
%   \newcommand{\cZ}{\cal{Z}}
%   \newtheorem{def}{Definition}[section]
%   ...

% add words to TeX's hyphenation exception list
%\hyphenation{author another created financial paper re-commend-ed Post-Script}

% declarations for front matter

\begin{document}

\begin{frontmatter}

%% Title, authors and addresses

%% use the tnoteref command within \title for footnotes;
%% use the tnotetext command for the associated footnote;
%% use the fnref command within \author or \address for footnotes;
%% use the fntext command for the associated footnote;
%% use the corref command within \author for corresponding author footnotes;
%% use the cortext command for the associated footnote;
%% use the ead command for the email address,
%% and the form \ead[url] for the home page:
%%
%% \title{Title\tnoteref{label1}}
%% \tnotetext[label1]{}
%% \author{Name\corref{cor1}\fnref{label2}}
%% \ead{email address}
%% \ead[url]{home page}
%% \fntext[label2]{}
%% \cortext[cor1]{}
%% \address{Address\fnref{label3}}
%% \fntext[label3]{}

\dochead{}
%% Use \dochead if there is an article header, e.g. \dochead{Short communication}

\title{The CMS Data Aggregation System}

%% use optional labels to link authors explicitly to addresses:
%% \author[label1,label2]{<author name>}
%% \address[label1]{<address>}
%% \address[label2]{<address>}

%\author[vkuznet]{Valentin Kuznetsov\corref{cor1}}
\author[vkuznet]{Valentin Kuznetsov}
\address[vkuznet]{Cornell University, Ithaca, New York, USA}
\ead{vkuznet@gmail.com}

\author[evans]{Dave Evans}
\address[evans]{Fermilab, Batavia, Illinois, USA}
\ead{evansde@fnal.gov}

\author[metson]{Simon Metson}
\address[metson]{Bristol University, Bristol, UK}
\ead{s.metson@bristol.ac.uk}

%\cortext[cor1]{Corresponding author}

\begin{abstract}
Metadata plays a significant role in large modern enterprises, 
research experiments and digital libraries where it comes from many different 
sources and is distributed in a variety of digital formats. 
It is organized and managed by constantly evolving software using 
both relational and non-relational data sources. Even though we can apply
an information retrieval approach to non-relational data sources,
we can't do so for relational ones, where information is accessed via
a pre-established set of data-services.

Here we discuss a new data aggregation system which consumes, 
indexes and delivers information from different relational and 
non-relational data sources to answer cross data-service queries 
and explore metadata associated with petabytes of experimental data. 
We combine the simplicity of keyword-based search with the precision of RDMS
under the new system. The aggregated information is collected from various sources,
allowing end-users to place dynamic queries, get precise answers and 
trigger information retrieval on demand. Based on the use cases of the CMS experiment, 
we have performed a set of detailed, large scale tests the results of which 
we present in this paper.
\end{abstract}

\begin{keyword}
%% keywords here, in the form: keyword \sep keyword
Meta-data\sep data aggregation\sep information discovery\sep HEP.

%% PACS codes here, in the form: \PACS code \sep code

%% MSC codes here, in the form: \MSC code \sep code
%% or \MSC[2008] code \sep code (2000 is the default)
\end{keyword}

\end{frontmatter}

%%
%% Start line numbering here if you want
%%
% \linenumbers

%% main text
\section{Introduction}
The European Organization for Nuclear Research, known as CERN, plays a leading
role in fundamental studies of physics. It is also known as a place where
many innovations in the area of computer science were developed, e.g. The World Wide Web.
Today, the Large Hadron Collider (LHC) at CERN is marking a new era of High Energy
Physics (HEP), promising to deliver a few PB of data each year. 
At this scale the ``information discovery'' within a heterogeneous, distributed 
environment becomes a key ingredient of successful data analysis.
The data and associated meta-data are produced in variety of forms and digital formats.
They are stored and retrieved from relational and non-relational data-sources, such as 
RDMS systems, document oriented databases, blogs, twikies, file systems,
customized applications, etc. Working in such an environment requires a lot 
of expertise and demands a straightforward, simple
way to look-up the desired information.
A well-known solutions, e.g. data-services, are tighten to a specific data 
sources and end-users are left with a manual task to bookkeep and 
relate information from them.

Here we present our work on Data Aggregation System (DAS) which provides
ability to search and aggregate information across different 
data-services. While designed for CMS High-Energy Physics 
experiment at LHC, the strategies and technology could be
used elsewhere.

The rest of this paper is organized as follows. 
In section \ref{RelatedWork} we discuss related work in a domain of 
keyword search over relational data sources.
Section \ref{DataModel} describes the CMS experiment and its data model. In section
\ref{DAS} we present the architecture of the DAS system, including discussion of its
various components. Finally, our results are summarized in section \ref{Results}.

\section{Related Work\label{RelatedWork}}
Even though the idea of keyword queries over the relational database
is not knew it is still under discussion in computer science domain.
A few alternative solutions have been proposed to address this issue
in last several years. In \cite{DBXplorer} the conjunctive keyword queries,
i.e. retrieval of documents that contain all query keywords, has
been discussed. It requires that all keywords are matched in a row from a
single table or joined tables. Authors of \cite{QueryAnswer} make a step forward
and have evaluated information around matched values. The results were presented as
customized templates to end-users. 
These and other related works, (see \cite{DBXplorer, QueryAnswer} 
for more references), are based on an approach that builds some 
kind of the graph over the underlying 
database and generates a new database out of the source database to perform the work.
The proximity of the results in those techniques are mostly confined to a 
text-based values stored in a database, since the usage of numerical ones 
is mostly useless without knowledge of the context of the input value. 
For example, typing a value 100 in the search input field can lead to 
plenty of inappropriate matches, e.g. row ids. In order to be
answered correctly it requires an additional keyword to clarify its meaning, 
which should be matched with particular entity (table.column) in the 
underlying database. Therefore the logical questions, e.g.
{\it find me the total number of files whose size more then 20 but less then 
100} are hard to answer precisely using approaches outlined in \cite{DBXplorer, QueryAnswer}, 
while it can be easily accomplished using SQL.

To address these limitations a keyword-based query language was introduced in \cite{DBS-QL}.
It used a set of pre-defined keys mapped to an underlying schema and shortest path
algorithm which employed foreign key constraints to build SQL queries. As a result
the question
{\it I'm looking for files who contain data taken on certain date and located at
particular site} was represented as simply as \cite{DBS-QL}
\begin{verbatim}
find file where date 2009-01-02 11:59 CET and site = T2
\end{verbatim}
without specifying any relational details.
This solution represents a common use case in the HEP community and provides
an intuitive mapping between the mental model of the end-users and 
the underlying database schema.

At the same time the data aggregation is commonly addressed in the context of
federated databases, see for example \cite{FedDB}. The data from different 
RDMS systems are stored into the federated DB where SQL queries can be used to 
search desired data. Apart from many internal obstacles, such as
schema, domain and naming conflicts, it is external issues that are often
a large barrier within a distributed, heterogeneous environment where
different security policies are in place. At the end the usage of SQL and/or
any other Query Language is required to search for data.

To avoid these limitations we resolved to build a keyword-search based system around
existing data-services whose nature and access policies are in place and well understood.
The key ingredients were the setup of the mapping and analytics services which provided
the necessary glue across existing API/notations and user queries. This approach 
allows us to perform dynamic data aggregation across various data-services, achieve
precise answers using keyword-based queries
and organize information retrieval on demand.

\section{CMS data model\label{DataModel}}
The Compact Muon Solenoid, (CMS) \cite{CMS} 
is one of the two general purpose particle physics detectors built on 
the Large Hadron Collider (LHC) at CERN in Switzerland and France. 
The experiment is designed to explore the frontiers of physics and provide physicists
with the ability to look at the conditions presented in the early stage of our Universe.
More then 3000 physicists from 183 institutions representing 38 countries 
are involved in the design, construction and maintenance of the experiment.

The CMS distributed computing and data model \cite{CMSDataModel} 
is designed to process and efficiently manage the few PBs of data expected each year
during normal operation of the LHC. The computing resources available to the CMS
collaboration are geographically distributed, 
interconnected via high throughput networks and operated by means 
of various Grid software infrastructure. The model is designed to
incorporate a broad variety of both mass data storage and data processing
resources into a single coherent system. CMS uses
multi-Tiered model, where specific tasks related to data taking such as
processing, archival and distribution, are assigned to each tier based
on the CMS data model. For example, the Tier-0 center at CERN is responsible
for archiving the data coming out from the detector, prompt first pass reconstruction
and data distribution to Tier-1 centers, located around the world.
The Tier-1 centers provide archival storage and CPU power for high
priority selection and re-reconstructions jobs and distribution points
to the Tier 2 resources.
The Tier-2 centers are dedicated for user analysis tasks and production of simulated data.

A broad variety of data-services have been designed and developed to
maintain detector and production operations, including detector
conditions databases, data bookkeeping services,
data transfer and job monitoring tasks. While the majority of these
services are located at CERN, they need to be accessed remotely, and
combined with other distributed services.

Once such a conglomerate of data-services start operating the obvious
question arises: how do we find the desired information across multiple data-services
in our distributed environment? Even though the individual data-services were designed
to answer specific questions about the data they serve, the ability to search and relate
information among them was remains a tedious human task. The growing amount of information
and desire to make cross-service queries led us to design and develop the new
Data Aggregation System.

\section{Data Aggregation System\label{DAS}}
The goal of the Data Aggregation System (DAS) was to fetch and aggregate meta-data 
on demand from existing CMS data-services under one umbrella by using their APIs, 
security policies, etc. In contrast, the approaches
discussed in \cite{DBXplorer, QueryAnswer, FedDB}
enforce a publication step of data into a dedicated database 
or central repository \cite{iRODS}. Therefore, quite often it was required
that data should be converted into common data structure/format in order to enable
aggregation, see \cite{OpenArchive}. In DAS the data are provided
by existing data-services and infrastructure, whose implementation,
data representation and security policies are left intact. Consequently,
DAS does not require a data preservation and transaction capabilities. 
Instead DAS can be treated as proxy or caching layer where information 
retrieval and aggregation are triggered dynamically by user queries. 
Below we discuss in details the DAS architecture.

\subsection{DAS architecture}
The design of the DAS system was based on previous studies of CMS Data 
Bookkeeping System (DBS) \cite{DBS, DBS07}. The DBS Data Discovery service
\cite{DD} was based on DBS Query Language (DBS-QL) \cite{DBS-QL}. It provided
a syntax similar to SQL, without specific knowledge of the underlying DB schema.
Quick adoption and wide usage of DBS-QL in CMS gave us confidence in the chosen approach 
and provided a basis for building the DAS system.

As was pointed out in \cite{Arms} the mixed content and 
mixed meta-data and meta-data consistency 
of the underlying services needs to be carefully
considered in the design of the system if it is to be a successful information 
discovery service.
Starting from this ground we are designed the DAS system as an
additional layer on top of the existing data-services
within CMS computing infrastructure by imposing the following set of requirements:
\begin{itemize}
\item support a keyword based search queries with ability to use conditional operators;
%\item pluggable interface to existing and up-coming data-service APIs;
\item use existing heterogeneous software environment and distributed nature of the data-services;
\item preserve access rules and security policies of individual data-services;
%\item be flexible to adjust to dynamic nature of data-service APIs, 
%due to continuously changing detector and software environments;
\item retrieve data on demand and aggregate them if necessary across
multiple data-services;
\item be transparent to data-services implementation and their data formats;
%\item support legacy data-services, applications and APIs.
\end{itemize}
\begin{figure}[htb]
\centering
\includegraphics[width=100mm]{DAS_Cache_and_Analytics.pdf}
\caption{
The DAS architecture diagram. It conists of cache server,
analytics and mapping databases. The data in a cache can be shared
across multiple queries. The information retrieval is triggered
on demand basis by invoking appropriate data-service APIs. The
optional DAS robots (Unix daemons) can be used to pre-fetch
the most populary query requests.
}
\label{DAS_cache}
\end{figure}

\noindent
The choice of existing RDMS as DAS back-end(s) were ruled out for several reasons. 
For instance, we didn't require any transactions and data persistency in DAS;
a dynamically typing of stored meta-data objects was one of the requirements, etc.
Those arguments forced us to look for alternative IT solutions.
Through the careful analysis of available options, such as a file-based and in memory caches, 
key-value databases, documented-oriented databases we made our choice in favor 
of the last technology. Among them we evaluated CouchDB \cite{CouchDB} and 
MongoDB \cite{MongoDB}. Both systems are ``schemaless'' databases providing
arbitrary document structure storage, replication and fail-over features. 
We made our decision in favor of MongoDB, due to its support of dynamic queries, 
full indexes, including inner objects and embedded arrays,
and auto-sharding. Our preliminary benchmarks have shown that it can sustain
the desired load and size capable of handling our meta-data information. We used 
document-oriented database, MongoDB, as a back-end for three 
main DAS components: Mapping, Analytics and DAS cache databases, 
which will be discussed next.

The DAS architecture is shown in Fig. \ref{DAS_cache}. It consists of
the core library with support of pluggable modules for data retrieval;
caching layer to store and aggregate results;
request service to handle user queries;
Mapping DB to keep information about data-service APIs, e.g.
notations and mapping to DAS keys;
Analytics DB for query analysis.

\subsection{DAS core\label{DAS_core}}
The DAS core library is responsible for data retrieval on demand. 
It has the following components: input query validator and parser,
query analyzer, workflow builder, execution engine and data-service 
plug-ins. The communication between client (web server or CLI) and 
DAS cache server was done via the REST \cite{REST} interface. 
Each input query was recorded into the Analytics DB for further 
analysis, and was decomposed into a set of conditions and/or selection keys.
First the results were looked up in a cache and delivered to the requestor
if superset of input conditions was found in analytics DB.\footnote{It is important
to note here that we didn't look-up results in a cache and rather
compared input query condition set to the records in analytics DB 
in order to understand if other API calls are covered the input request.}
Otherwise, the appropriate workflow, a set of API calls, was constructed
for the execution engine. It invoked the API calls via data-service
plug-ins and stored results into the cache. At this step we also
performed necessary data aggregation for records with similar key-value
pairs. The data-service plug-in implementations are similar and their 
logic is largely abstracted, apart from a few extentions such as specific
security modules for data-services accessible via private network.
Finally, the results were retrieved from the cache and delivered to the user.

\subsection{DAS Mapping DB\label{MappingDB}}
The Mapping DB is used to track services known to DAS and how to interact with them.
%The Mapping DB represents DAS {\it registration} service. 
It collects
information about data-service APIs, translates API input/output
parameters into DAS keys and contains mapping between DAS records
and their UI representation.

On identifying API's that are accessible to DAS we store
their names, input parameters and associated DAS keys
into the Mapping DB. For example
\begin{verbatim}
{"api":{"params":{"name":""}, "name":"CMSNametoSE"}, "system":"sitedb",
 "api2das": [{"pattern": "re.compile('^T[0-3]_')", 
              "api_param": "name", "das_key": "site"}], 
 "daskeys": [{"pattern": "re.compile('^T[0-3]_')", 
              "key": "site", "map": "site.name"}]}
\end{verbatim}
represents DAS mapping record.\footnote{
The data-service URL's are kept in the configuration file.
}
This record can be read as following:
the \verb+CMSNametoSE+ API provided by 
\verb+sitedb+ data-service accepts \verb+name+ as input parameter. It
confirms regular expression pattern, \verb+re.compile('^T[0-3]_')+, and
corresponds to \verb+site+ DAS key. The \verb+site+ key
maps into \verb+site.name+ within DAS record.

Therefore the Mapping DB identifies a set of pre-defined DAS keys, 
e.g. \verb+site, file, run+, etc., to be used by end-users in their queries.
Each key is mapped into set of API calls, whose output is retrieved,
cached and aggregated as required.

\subsection{DAS Analytics DB}
The DAS Analytics DB collects information on user requests against the system.
Each request that DAS handles is recorded.
The Analytics DB tracks:
\begin{itemize}
\item The users request
\item How DAS has mapped the request onto data service API's
\item The time taken by the remote data services to process the API calls
\end{itemize}
By recording this information we can plan pre-fetch strategies for common queries,
identify issues in remote data services and cross check that DAS resolves requests
in a deterministic manner.

% Upon its
%decomposition into set of selection keys and conditions we also recorded
%API name and input parameters into the Analytics DB and 
%kept updating their counter information for repeated calls. 
%This allowed to keep track of frequency of API calls and establish 
%pre-fetch strategies for most common queries.

\subsection{DAS caching system}
The DAS caching system is used to dynamically fetch
and aggregate data upon user requests. We use the following 
algorithm (code snippets are written in Python \cite{Python}):
\begin{enumerate}[1.]
\item Decompose query into set of conditions (\verb+conds+) 
and selection keys (\verb+skeys+).
\item Check cache for query hash. If found, look-up results from the cache,
otherwise proceed to the next step.
\item Identify list of APIs which can serve requested data
\begin{verbatim}
apilist = [r for r in mapping_db(skeys, conds)]
\end{verbatim}
Here \verb+r+ represents a triplet \verb+url, api, args+ and 
\verb+mappind_db+ defines translation between input query and data-service
APIs which can delivery the data.
\item Iterate over APIs
\begin{verbatim}
for url, api, args in apilist:
    if  superset_in_analytics(api, args):
        continue
    for row in parser(get_data(url, api, args)):
        yield row
\end{verbatim}
Here \verb+superset_in_analytics+ checks if provided \verb+api, args+
are subset of already registered in Analytics DB api call(s). The 
\verb+get_data+ and \verb+parser+ are define methods to fetch and parse
data, respectively.
\item Perform aggregation by iterating over records from the previous step
\begin{verbatim}
if cache_has(daskey):
   existing_record = get_from_cache(daskey)
   merge(row, existing_record)
   insert_into_cache(row)
   delete_from_cache(existing_record)
else:
   insert_into_cache(row)
\end{verbatim}
The \verb+daskey+
is a key which identifies the record defined in Mapping DB, e.g. \verb+file.name+
for \verb+file+ record.
We did use a bulk insert operations at this step to increase throughput. 
\item Get results from the cache
\end{enumerate}
We also have an option to run independent set of daemons, DAS robots, who
monitor Analytics DB for most popular requests and pre-fetch the most
common data into the cache.

\subsection{DAS queries}
All DAS queries are expressed in a free text-based form, either as a 
set of keywords or key-value pairs, where a pair can represent
condition, e.g. \verb+key > value+. We use a pre-defined set of keys registered 
in DAS to describe common entities of our data. For example, \verb+dataset+, 
\verb+file+, \verb+block+, \verb+run+, \verb+event+, etc. Their choice
is based on the analysis of daily work operations of our users, see \cite{DBS07}.
Each key represents a DAS record. Every record is stored into the cache
in JSON format. Due to the ``schemaless'' nature of underlying back-end (MongoDB)
we are able to store DAS records of arbitrary structure, e.g.
dictionary, lists, key-value pairs, etc. Therefore every DAS key has a 
set of attributes representing JSON structure. For instance a document
\begin{verbatim}
{"block": {"name": "abc", "size": 1, "replica": [...]}}
\end{verbatim}
is associated with a \verb+block+ key who has the following attributes
\verb+name+, \verb+size+, \verb+replica+. The set of attributes is based 
on a set of meta-data information stored in a record and therefore
a nested structures are allowed, e.g. \verb+block.replica.size+. Our users 
are able to specify a context of their queries by using appropriate 
attribute(s). For example, query \verb+block+ represents search for all
block records rather then ``block'' string, while query
\verb+block.name=abc*+ instructs the system to look-up appropriate blocks
with matched names. The numerical values, conditions, grouping are
easy to express too, e.g. \verb+run > 10 and run < 100+ query 
is used to get all run records for provided range of run numbers. 
To complement the strength of the queries, we do plan to index 
dynamically DAS records and add a keyword-based 
search within record content later.

All input queries are transformed into MongoDB query syntax who is itself
a JSON record. It contains a set of selection 
keys (\verb+fields+) and conditions (\verb+spec+). For example, a query
\verb+dataset, file=abc, run>10+
is represented as
\begin{verbatim}
{"fields": ["dataset"], "spec": {"file": "abc", "run": {"$gt": 10}}}
\end{verbatim}
Therefore users queries are stored into Analytics DB similar to other DAS 
records and used for further analysis.

\subsection{Data-services and aggregation}
As was shown in section \ref{MappingDB} the DAS Mapping DB became an authorative
source of information about all data-services participating in DAS.
Based on the information stored in Mapping DB we were able to 
retrieve appropriate data from different data-services, re-map them into 
DAS notations and aggregate them on demand. For instance, the Data 
Bookkeeping System (DBS) \cite{DBS} and data location system (PhEDEx) \cite{PhEDEx}
provide information about CMS blocks. In former case, DBS 
stored information about blocks and their relationships to other 
physics objects, such as dataset and files, while PhEDEx kept
information about its location and replicas.

>>>>

\documentclass[a4paper]{jpconf}
\usepackage{graphicx}
\begin{document}
\title{The CMS Data Aggregation System}

\author{Valentin Kuznetsov}
\address{Cornell University, Ithaca, New York, USA}
\ead{vkuznet@gmail.com}

\author{Dave Evans}
\address{Fermilab, Batavia, Illinois, USA}
\ead{evansde@fnal.gov}

\author{Simon Matson}
\address{Bristol University, Bristol, UK}
\ead{s.metson@bristol.ac.uk}

%In a large modern enterprise, information is almost inevitably distributed among several database management systems. Despite considerable attention from the research community, relatively few commercial systems have attempted to address this issue. This article describes the technology that enables clients of IBM's federated database engine to access and integrate the data and specialized computational capabilities of a wide range of relational and non­relational data sources.

\begin{abstract}
A metadata plays significant role in a large modern enterprises, research experiments,
digital libraries where it comes from different sources and distributed in a 
variety of forms and digital formats. It is organized and managed by constantly
evolving software using both relational and nonrelational data sources. There is
a big demand to access information from multiple sources.
Here we discuss a new data aggregation system which deliver and cache information 
from different relational and nonrelational data sources on a concrete example 
of large scale High-Energy physics experiment.
\end{abstract}

\newpage

\section{Introduction}
The CERN, the European Organization for Nuclear Research, plays a leading
role in fundamental studies of physics. But apart from scientific community 
it is known as a place where World Wide Web was born. At that time the 
information look-up among hyperlinks represents a certain challenge for scientists.
Today, the Large Hadron Collider (LHC) at CERN is marking a new era of High Energy
Physics (HEP), promising to deliver a few PB of data each year. Scientists are
facing another set of problems from distributed computing to multi-dimentional
view on information retrieval. One of the aspect of such research is an efficient
and at the same time precise look-up of produced meta-data, which comes in variety 
of forms and digital formats. As was pointed out in \cite{Amr} a mixed content and 
mixed metadata and metadata consistency should be considered as a whole in design 
of the system to successful information discovery.

The CMS, the Compact Muon Solenoid experiment, operated at LHC,
represents an heterogeneous environment of distributing computing, relational and
nonrelational data sources where we face this problem. The computing resources
are distributed among almost 40 countries, 183 institutions and more then 3000 physicists.
Broad variety of RDMS systems at different data centers collect a various
metadata information, including data location and transfer, detector conditions,
calibrations constants, simulation information, etc. Moreover the development
of those components were done in parallel by different group and technology
tools. Therefore a wide spectrum of data services and
data formats represents a problem of information discovery.

\section{Related Work}
Even though the idea of querying relation databases via keyword based search
algorithm is not knew it is still under significant activity in computer
sceince domain. A few
alternative solutions has been proposed to address this issue. On one
side, the federated DB \cite{FedDB} unify data coming from different
RDMS into another DB where SQL queries can be placed to search desired
data. While another approaches of querying relational
DB(s) using keyword search algorithms has been proposed recently, see
\cite{DBXplorer, QueryAnswer}.
In former case, you still face with understanding of the underlying schema and
imposing relational conditions in your query, while queries can be
expressed via simple keyword based search, even though exact match
of provided keywords is expected. Although keyword based search is
intuitive and easy to use, with respect to HEP domain it is not sufficient. 
For example, how to deal with numbers? What a single number means in a user input?
The row ID, the value of the field, the sum of files or their size? In other
words a further step is required to understand up-coming request without
sacrifies the output results match. Our users can be classified as proficient
users who understand semantics of the problem and want precise answers
for their queries. They are more interesting in precise answers to their
questions rather ranking output.

In \cite{DBS-QL} a simple, intuitive and flexible query language was introduced 
on top of the CMS data-bookkeeping system. It represented a power of SQL while
hiding underlying relational schema from the end users. As a results
a human questions were intuitively mapped into simple queries. For example,
a question
{\it I'm looking for files who contain data taken on certain date and located at
particular site} was represented as simple as
\begin{verbatim}
find file where date 2009-01-02 11:59 CET and site = T2
\end{verbatim}
We wanted to expand this approach and apply such queries across multiple
data services and different DB back-ends.

Here we discuss a new system, the Data Aggregation System (DAS),
developed in CMS collaboration to address this issue. We start with
discussion of CMS data model and existing data-services, see \ref{DataModel}. 
In section \ref{DAS}
we provides a detailed description of underlying design and components
used in DAS system. Our results are summarized in section \ref{Results}.

\section{CMS data model\label{DataModel}}
\section{Data Aggregation System\label{DAS}}
Data Aggregation System shown in Figure \ref{DAS_arch} and
consists out of several components:
\begin{figure}[htb]
\centering
\includegraphics[width=150mm]{DAS_architecture.pdf}
\caption{
DAS architecture.
}
\label{DAS_arch}
\end{figure}

\begin{itemize}
\item DAS core system which is support pluggable modules to retrieve data from
CMS data-services, e.g. Data Bookkeeping System (DBS), data location service (Phedex),
etc.;
\item DAS caching system which has dual purpose: to store ``raw'' results coming
out from data-services, a raw-cache, and aggregated results produced by DAS core into
hot-cache;
\item DAS request service who collect and place into queue user requests
\item DAS mapping DB to store mapping between data-services and DAS notations and
data-representation;
\item DAS analytics DB to collect query statistics and establish pre-fetching 
strategies.
\end{itemize}
We will discuss each of those components in more details in up-coming sub-sections.
The DAS supports a pluggable approach to add a new CMS sub-system into a game by
identifying appropriate APIs, input and output parameters and keys, data-service
format and mapping between data-service, DAS and other sub-system. For instance,
the Data Bookkeeping System (DBS) and data location system (Phedex) store information about
a file. In former case, a file name was stored as well as its relations to other physics 
objects, such as run and luminosity, while in later case the Phedex store information 
about file location. When user specify a file as their selection key, based on
information store in DAS mapping DB we were able to identify appropriate set of
data-service APIs, look-up data, re-format their output to DAS notations and
store it into DAS raw-cache. Even though we support a data retrieval on demand
basis we also store user queries into DAS analytics DB to identify
a common patterns. Based on this information we organized a pre-fetching
populators. Once data were coming to raw-cache an additional step of
aggregation was applied. If other data with identical keys were presented in
a raw-cache we merge it them together. For instance upon the following query
$find file where ... $ a file information from DBS and Phedex systems were
retrieved. Since both data-services output contained file name as a common
key, their output where merged and final result was stored into raw-cache.
This merged (aggregated) results were shipped back to client and stored into
hot-cache (memcached) to speed up a DAS web client application. Each DAS data
object contains a standard DAS header, which identify look-up time,
request url, API, parameters and method as well as response version, expiration
time and checksum of the output.

\subsection{DAS Query Language}
Upon success of DBS-QL \ref{DBS-QL} we adopted its syntax to DAS with a few minor
changes. Here we briefly recall its syntax:
\begin{equation}
action\,\,\,
key_1,\, key_2,\, ...\,\,\, where\,\,\,
\langle key\rangle\, 
\langle op\rangle\, 
\langle value\rangle \,\,\, and|or ...
\label{QL_syntax}
\end{equation}
Each DAS QL expression starts with special keyword indicating an action such as 
$find,\, plot,\, summary$, followed by a set of selection keys. The action identify
the behavior of the query, find to look-up data, plot to plot found data, summary to
apply a summary view over provided selection keys. An optional conditions can be 
specified via $where$ keyword following by condition expression.
Each condition expression represented via $key\,\,\, operator\,\,\, value$ triplet, where
$key$ is DAS selection key, $operator$ is conditional operator, 
$=,\, >=,\, <,\, in,\, between$ and $value$ is a value applied to selection key.
We also support brackets as well as $and$ and $or$ logical operators.
The DAS selection and condition keys were identified based on analysis 
of terminology used by our users. For instance, $file$ was used to either
select file name or used in condition to identify file name provided by user.
Here is a simple example of DAS QL expression:
$$
find\,\,\, file,\, run\,\,\, where\,\,\, site=T1\_CH\_CERN
$$
which represents a natural language question: 
{\it find files and runs located at T1\_CH\_CERN site}.

\subsection{DAS Mapping DB}
As been discussed above we used a DAS mapping DB to map data-service API into
DAS as well as map provide a relations between different data-services.
Upon identification of APIs which participate in DAS activity 
we store a mapping between data-service notations
into DAS mapping DB. This allows to map user input query parameters into
data-service API metrics as well as map back data-service output into DAS names.
For instance, the DBS system used $logical\_file\_name$ notation, while in Phedex
the same entity was named as $lfn$. Since both of them were referred to
logical file name, we map both notations into $file$ notation in DAS.\footnote{We
also deal with physical file names, but still used $file$ to refer to them in
DAS notation, since simple regular expressions were used to identify $file$
meaning from the provided value.}
For each service we provide a mapping from its own data-format (JSON, XML, etc.) into
common data-format used in DAS.\footnote{We used JSON as a DAS data format}
We also store a mapping between DAS keywords and data-service API names to
establish appropriate calls to data-service upon provided DAS query. For example,
when user specifies $file$ in their DAS query, we invoke DBS listFiles and Phedex
fileReplicas APIs. Since all mapping kept in RDMS we were able to abstract the 
DAS core library calls from hardcoded data-service API calls and make it
transparent to any participating data-services. Here is a logic of DAS core
library:
\begin{itemize}
\item parse input DAS query
\item map provided selection and condition keys into data-service APIs and
input parameters
\item invoke data-service APIs
\item parse output results by mapping output notation into DAS keys
\item store and aggregate them into DAS raw-cache
\end{itemize}

\subsection{DAS Analytics DB}

\subsection{Caching system}
The DAS caching system was divided into two different caches: raw-cache and hot-cache.
The purpose of raw-cache was to store results coming from data services for
as long as specified in data-service response header. While the purpose of hot-cache
was to keep final DAS results corresponding to provided user query. We use a populator
services (jobs) to feed raw-cache with highly requested queries. Here we will only
discuss the raw-cache organization. 
\begin{figure}[htb]
\centering
\includegraphics[width=150mm]{DAS_cache.pdf}
\caption{
DAS cache system.
}
\label{DAS_cache}
\end{figure}
As it shown in figure \ref{DAS_cache} we setup
dedicated DAS cache request server to manage user requests. Each request placed into
queue who was used by workers to spawn a data-service request. 
This allows us to utilze a multi-core technology and place a workers at each 
core.\footnote{A number of workers per core were adjustable to satisfy requests load.}
Each worker talk to DAS raw-cache and either stored mapped results over there or merge
them with existing data object in DAS cache. We evaluated several technologies
as a choice for DAS cache back-end, including relational MySQL DB \cite{MySQL} 
and object oriented CouchDB \cite{CouchDB} and MongoDB \cite{MongoDB}.
Each of this technologies had its own pros and cons. For instance usage
of MySQL provided a SQL queries (as in federated DBs, see \cite{FedDB}), but
required auto-schema generation on a fly.\footnote{We did not apply any
restriction to DAS QL and did not know in advance what our users will be
querying.} While, using CouchDB and MongoDB, we were able to store
data-service output naturally as data objects into DB. Finally we decided
to use MongoDB due to its flexible and
powerful querying syntax, which were easy to map into DAS QL expressions.

\subsection{REST interface}
We designed DAS interface to be fully REST complaint \cite{REST}.
All requests placed into DAS cache server were done via GET/PUT/POST/DELETE methods.
DAS clients initiate their request via GET method. An appropriate message
with either resulting data or their absense was send back to client.
Clients were able to decided upon received message how to proceed, either
request data via POST/PUT methods, DELETE data in cache or initiate another
GET request. The REST complaince help us to treat our clients, user quyering
DAS system via web browser, other CMS applications and command line tools
in identical manner. Also, the implementation of pre-fetch populator
tools become as easy and identical as any other clients.

\section{Results\label{Results}}

\section{Summary}

\section{Acknowledgements}
